---
title: "tug-of-war"
author: "M. H. Tessler"
date: "November 1, 2016"
output: html_document
---

```{r helpers, echo = F}
library(knitr)
knitr::opts_chunk$set(fig.crop = F,echo=T, 
                      warning=F, cache=F, 
                      message=F, sanitize = T)

library(rwebppl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(coda)
library(jsonlite)

# # to get the maximum a posteriori value [MAP] (from samples)
# estimate_mode <- function(s) {
#   d <- density(s)
#   return(d$x[which.max(d$y)])
# }
# # upper bound of 95% credible interval
# hdi_upper<- function(s){
#   m <- HPDinterval(mcmc(s))
#   return(m["var1","upper"])
# }
# # lower bound of 95% credible interval
# hdi_lower<- function(s){
#   m <- HPDinterval(mcmc(s))
#   return(m["var1","lower"])
# }
```

## Tug of war model

Possible questions: 

1. "What is the strength of A?"
2. "Was A being lazy in match N?"
3. "Who will win: A vs. B?""

```{r towModel}
towModel <- '
// unpack data from R
var lazinessPrior = datafromR["lazinessPrior"][0]
var lazyPulling = datafromR["lazyPulling"][0]
var matchInfo = datafromR["matchInfo"][0]

var tugOfWarModel = function(){

    var strength = mem(function(person) {
      return gaussian(50, 6)
    })
    
    var lazy = mem(function(person, match) {
      return flip(lazinessPrior)
    })
    
    var pulling = function(person, match) {
      return lazy(person, match) ? strength(person) * 0.5 : strength(person)
    }
    
    var totalPulling = function(team, match) {
      return sum(map(function(person) {
        return pulling(person, match)
      }, team))
    }
    
    var winner = function(team1, team2, match) {
      return totalPulling(team1, match) > totalPulling(team2, match) ? team1 : team2
    }
    
    var beat = function(team1,team2, match){
      return winner(team1,team2, match) == team1
    }
    
    var conditions = function(i, len) {
      if (i < len) {
        if (matchInfo.games[i].winner == 1) {
          condition(beat(matchInfo.games[i].team1,
                        matchInfo.games[i].team2,
                        i + 1));
        } else if (matchInfo.games[i].winner == 2) {
          condition(beat(matchInfo.games[i].team2,
                        matchInfo.games[i].team1,
                        i + 1));
        }
        conditions(i + 1, len);
      }
    }

    conditions(0, matchInfo.games.length);


'
```

Define match information (the only information that's used in the model above is the winner/loser information, though the other information would be used in data analysis models).

Also, define the lazinessPrior and lazyPulling variables, package them up into `dataToWebPPL`, and pass it to the WebPPL model using the `data` and `data_var` options. (?webppl for more details)

```{r runTowModel}
runModel <- function(scenarioID, shouldPlot) {
  
  jsonFile = "stim.json"
  stim <- fromJSON("stim.json")
  scene = stim$scenarios[(scenarioID+1),]
  
  retObject = 'return { '
  
  
  
  for (i in 1:length(scene$questions[[1]])) {
    elementToAdd = ""
    if (scene$questions[[1]][i] == 0) {
      elementToAdd = paste0("'strength", scene$subjects[[1]][i, 1], "': strength(", scene$subjects[[1]][i, 1], ")")
    } else if (scene$questions[[1]][i] == 1) {
      elementToAdd = paste0("'laziness", scene$subjects[[1]][i, 1], scene$subjects[[1]][i, 2], "': lazy(", scene$subjects[[1]][i, 1], ", ", scene$subjects[[1]][i, 2], ")")
    }
  
    if (i < length(scene$questions[[1]])) {
      elementToAdd = paste0(elementToAdd, ", ")
    }
    retObject = paste0(retObject, elementToAdd)
  }
  
  retObject = paste0(retObject, "}}")
  
  towModel = paste0(towModel, retObject)
  
  dataToWebPPL = list(lazinessPrior = 0.3, 
                      lazyPulling = 0.5,
                      matchInfo = scene)
  
  rs <- webppl(program_code = towModel, 
               model_var = "tugOfWarModel",
    inference_opts = list(method = 'rejection', samples = 1000),
    data = dataToWebPPL,
    data_var = "datafromR",
    output_format = "samples"
  )
  
  # Plotting the results
  rs.tidy <- rs %>%
    gather(key, val)
  
  rs.tidy$val <- as.double(rs.tidy$val)
  
  if (shouldPlot) {
    
    print(mean(rs$strength1))
    
    ggplot(rs.tidy, aes( x = val ))+
    geom_histogram() +
    facet_wrap(~key, scales = 'free')
    
  } else {
    return(mean(rs$strength1))
  }
  
  
}

df <- data.frame(strength=integer())

getMeans = function(n) {
  for (i in n) {
    m = runModel(i, FALSE)
    df[i,] <- c(m) 
  }
  
  df <- transform(df, strength.rank=ave(strength, FUN=function(x) {rank(x)}))
  
  ggplot(df, aes(x = strength.rank, y = strength))+
    geom_bar(stat='identity')+
    geom_text(data=df, aes(label=rownames(df)))+
    coord_cartesian(ylim=c(45, 55))
}

#You can run the model on just one scenario in stim by calling runModel(id, TRUE)
  
runModel(29, TRUE)
#getMeans(21:30)
```