---
title: "tug-of-war"
author: "M. H. Tessler"
date: "November 1, 2016"
output: html_document
---

```{r helpers, echo = F}
library(knitr)
knitr::opts_chunk$set(fig.crop = F,echo=T, 
                      warning=F, cache=F, 
                      message=F, sanitize = T)

library(rwebppl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(coda)
library(jsonlite)
rm(list=ls())

# # to get the maximum a posteriori value [MAP] (from samples)
# estimate_mode <- function(s) {
#   d <- density(s)
#   return(d$x[which.max(d$y)])
# }
# # upper bound of 95% credible interval
# hdi_upper<- function(s){
#   m <- HPDinterval(mcmc(s))
#   return(m["var1","upper"])
# }
# # lower bound of 95% credible interval
# hdi_lower<- function(s){
#   m <- HPDinterval(mcmc(s))
#   return(m["var1","lower"])
# }
```

## Tug of war model

Possible questions: 

1. "What is the strength of A?"
2. "Was A being lazy in match N?"
3. "Who will win: A vs. B?""

```{r towModel}
towModel <- '

// unpack data from R
var lazinessPrior = datafromR["lazinessPrior"][0]
var lazyPulling = datafromR["lazyPulling"][0]
var matchInfo = datafromR["matchInfo"][0]
var closeThreshold = datafromR["closeThreshold"][0]
var whoIs = datafromR["whodata"][0]





// helper function
var exp = function(x){return Math.exp(x)}

// for discretization
var binParam = 5; 

var gaussianParams = {mu: 50, sigma: 10};

// calculate the range in pre-defined steps;
// these values correspond to possible heights
var stateVals = _.range(gaussianParams.mu - 3 * gaussianParams.sigma,
                        gaussianParams.mu + 3 * gaussianParams.sigma,
                        gaussianParams.sigma/binParam)

// for each possible height, calculate its probability of occurrence
var stateProbs = cache(function(stateParams){
  return map(function(s){
    exp(Gaussian(stateParams).score(s))
  }, stateVals)
});
///

var DiscreteGaussian = cache(function(stateParams) {
  return Infer({
    model: function(){ return categorical({vs: stateVals, ps: stateProbs(stateParams)}) }
  })
});

// generate the uniform threshold prior
var thresholdBins = cache(function(utt, stateSupport){
  return map(function(x){
    return utt == "strong" ? x - (1/(binParam*2)) : x + (1/(binParam*2));
  }, sort(stateSupport))
})

var thresholdPrior = cache(function(utt, stateSupport){
  return Infer({
    model: function() { return uniformDraw(thresholdBins(utt, stateSupport)) }
  });
});

var somethingCost = 5
var cost = {
  "strong": somethingCost,
  "weak": somethingCost,
  "": 0
};

var UtterancePrior = cache(function(utt) {
  var utterances =  [utt].concat([""])
  var uttProbs = map(function(u) {return exp(-cost[u]) }, utterances);
  return Categorical({ps: uttProbs, vs: utterances})
});


// meaning function for utterances
var meaning = function(utterance, state, threshold) {
  utterance == "strong" ? state > threshold ? flip(0.9999) : flip(0.0001) :
  utterance == "weak" ? state < threshold ? flip(0.9999) : flip(0.0001) :
  true
}

// set sepeaker optimality
var alpha = 10;
var statePrior = DiscreteGaussian(gaussianParams)

var literalListener = cache(
  function(utterance, threshold) {
    Infer({model: function(){
      var state = sample(statePrior);
      var m = meaning(utterance, state, threshold);
      condition(m);
      return state;
    }})
  }, 10000 // limit cache size
)

var speaker1 = cache(
  function(state, threshold, actualUtt) {
    Infer({model: function(){
      var utterance = sample(UtterancePrior(actualUtt));
      var L0 = literalListener(utterance, threshold);
      factor( alpha * L0.score(state) );
      return utterance;
    }})
  }, 10000 // limit cache size
)

var pragmaticListener = cache(function(utterance) {
  Infer({model: function(){
    var state = sample(statePrior);
    var threshold = sample(thresholdPrior(utterance, statePrior.support()))

    var S1 = speaker1(state, threshold, utterance);
    observe(S1, utterance);
    return state 
  }})
}, 10000)

var strengthDistribution = {
  strong:  pragmaticListener("strong"),
  weak: pragmaticListener("weak"),
  normal: Gaussian({mu: 50, sigma: 10})
}



var tugOfWarModel = function(){

  var getWhoIsWhat = function(i, whoIsWhat) {
    return whoIsWhat[i-1]["is"];
  }

  var whoIsWhat = {
    1: getWhoIsWhat(1, whoIs),
    2: getWhoIsWhat(2, whoIs)
  }

  var strength = mem(function(person) {
    var isSpecial = whoIsWhat[person]
    return sample(strengthDistribution[isSpecial ? isSpecial : "normal"])
  })


    var uuid = function() {
      var s4 = function() {
        return (Math.floor((1 + Math.random()) * 0x100)
                .toString(16)
                .substring(1));
      }
      return s4()
    }
    
    var generateGame = function(player, winnerTeamIndex){
      var team1size = uniformDraw([1,2,3])
      var team2size = flip(0.7) ? team1size : uniformDraw([1,2,3])
      return {
        "team1": [player].concat(repeat(team1size - 1, uuid)),
        "team2": repeat(team2size, uuid),
        "winner": winnerTeamIndex
      }
    }
    
    var wonNPreviousGames = function(player, nWon, nTotal){
      return repeat(nWon, function(){ generateGame(player, 1) }).concat(
        repeat(nTotal - nWon, function(){ generateGame(player, 2) })
      )
    }
    
    var lazy = mem(function(person, match) {
      return flip(lazinessPrior)
    })

    var lazyRandom = function() {
      return flip(lazinessPrior)
    }
    
    var pulling = function(person, match) {
      return lazy(person, match) ? strength(person) * lazyPulling : strength(person)
    }
    
    var totalPulling = function(team, match) {
      return sum(map(function(person) {
        return pulling(person, match)
      }, team))
    }
    
    var deltaTeamStrength = function(team1, team2, match){
    return totalPulling(team1, match) - totalPulling(team2, match)
    }

    var isClose = function(team1, team2, match){
    return Math.abs(deltaTeamStrength(team1, team2, match)) < closeThreshold
    }  

    var winner = function(team1, team2, match) {
    return deltaTeamStrength(team1, team2, match) > 0 ? team1 : team2
    }
    
    var beat = function(team1,team2, match) {
      return winner(team1,team2, match) == team1
    }

    var wouldBeat = function(player1, player2) {
      var p1 = lazyRandom() ? strength(player1) * lazyPulling : strength(player1)
      var p2 = lazyRandom() ? strength(player2) * lazyPulling : strength(player2)
      
      return p1 > p2
    }
    
  var conditions = function(i, listOfGames) {
    var len = listOfGames.length;
    if (i < len) {
      if (listOfGames[i].winner == 1) {
        condition(beat(listOfGames[i].team1,
                       listOfGames[i].team2,
                       i + 1));
      } else if (listOfGames[i].winner == 2) {
        condition(beat(listOfGames[i].team2,
                       listOfGames[i].team1,
                       i + 1));
      }
      conditions(i + 1, listOfGames);
    }
     }

    conditions(0, matchInfo.games);

    var condition_wonNPrev = function(i, listOfPrevs) {
      var len = listOfPrevs.length;
        if (i < len) {
          conditions(0, wonNPreviousGames(listOfPrevs[i].player, 
                                  listOfPrevs[i].nWon, 
                                  listOfPrevs[i].total));
          condition_wonNPrev(i + 1, listOfPrevs);
        }
    }
    
    condition_wonNPrev(0, matchInfo.wonngames);

    
  var conditions_wasClose = function(i, listOfClose, listOfGames) {
    var len = listOfClose.length;
    if (i < len) {
      condition(isClose(listOfGames[listOfClose[i]-1].team1,
                       listOfGames[listOfClose[i]-1].team2,
                       listOfClose[i]));
      conditions_wasClose(i + 1, listOfClose, listOfGames);
    }
  }

  conditions_wasClose(0, matchInfo.closegames, matchInfo.games)

'
```

Define match information (the only information that's used in the model above is the winner/loser information, though the other information would be used in data analysis models).

Also, define the lazinessPrior and lazyPulling variables, package them up into `dataToWebPPL`, and pass it to the WebPPL model using the `data` and `data_var` options. (?webppl for more details)

```{r runTowModel}
scenarioID = 1
shouldPlot = TRUE
#runModel <- function(scenarioID, shouldPlot) {
  
  stim <- fromJSON("../../javascript/experiment_3/static/json/n_exp3_stim.json")
  scene = stim$scenarios[(scenarioID+1),]
  whoiswhat = stim$scenarios[(scenarioID+1),]$whoiswhat
  
  retObject = 'return { '
  
  
  for (i in 1:length(scene$questions[[1]])) {
    elementToAdd = ""
    if (scene$questions[[1]][i] == 0) {
      elementToAdd = paste0("'strength", scene$subjects[[1]][i, 1], "': strength(", scene$subjects[[1]][i, 1], ")")
    } else if (scene$questions[[1]][i] == 1) {
      elementToAdd = paste0("'laziness", scene$subjects[[1]][i, 1], scene$subjects[[1]][i, 2], "': lazy(", scene$subjects[[1]][i, 1], ", ", scene$subjects[[1]][i, 2], ")")
    } else if (scene$questions[[1]][i] == 2) {
      elementToAdd = paste0("'beat", scene$subjects[[1]][i, 1], scene$subjects[[1]][i, 2], "': wouldBeat(", scene$subjects[[1]][i, 1], ", ", scene$subjects[[1]][i, 2], ", 100)")
    }
  
    if (i < length(scene$questions[[1]])) {
      elementToAdd = paste0(elementToAdd, ", ")
    }
    retObject = paste0(retObject, elementToAdd)
  }
  
  retObject = paste0(retObject, "}}")
  
  newModel = paste0(towModel, retObject)
  
  dataToWebPPL = list(lazinessPrior = 0.3, 
                      lazyPulling = 0.5,
                      closeThreshold = 10,
                      whodata = whoiswhat,
                      matchInfo = scene)
  
  rs <- webppl(program_code = newModel, 
               model_var = "tugOfWarModel",
    inference_opts = list(method = 'rejection', samples = 100),
    data = dataToWebPPL,
    data_var = "datafromR"
  )
  
  # Plotting the results
  rs.tidy <- rs %>%
    gather(key, val)
  
  rs.tidy$val <- as.double(rs.tidy$val)
  
  if (shouldPlot) {
    
    for (i in names(rs)) {
      print(mean(rs[[i]]))
    }
    
    ggplot(rs.tidy, aes( x = val ))+
    geom_histogram() +
    facet_wrap(~key, scales = 'free') 
    
  } else {
    #return(mean(rs$strength1))
  }
  
  
#}

df <- data.frame(strength=integer())

#You can run the model on just one scenario in stim by calling runModel(id, TRUE)
  
#runModel(30, TRUE)
#getMeans(21:30)
```

```{r getModelPredictions}
stim <- fromJSON("../../javascript/experiment_3/static/json/n_exp3_stim.json")

firstGame = 4
# lastGame = 3
lastGame = 12

df.strength <- data.frame(id=character(), rating=double(), stringsAsFactors = FALSE)
df.lazy <- data.frame(id=character(), rating=double(), stringsAsFactors = FALSE)
df.wouldwin <- data.frame(id=character(), rating=double(), stringsAsFactors = FALSE)

sIndex = 0
lIndex = 0
wIndex = 0
cIndex = 0

for (x in (firstGame + 1):(lastGame + 1)) {
  scene = stim$scenarios[x,]
  
  for (i in 1:length(scene$questions[[1]])) {
    retObject = 'return { '
    elementToAdd = ""
    ide = "0"
    if (scene$questions[[1]][i] == 0) {
      sIndex = sIndex + 1
      elementToAdd = paste0("'strength", scene$subjects[[1]][i, 1], "': strength(", scene$subjects[[1]][i, 1], ")")
      ide = paste0(x-1, "-", scene$subjects[[1]][i, 1])
    } else if (scene$questions[[1]][i] == 1) {
      lIndex = lIndex + 1
      elementToAdd = paste0("'laziness", scene$subjects[[1]][i, 1], scene$subjects[[1]][i, 2], "': lazy(", scene$subjects[[1]][i, 1], ", ", scene$subjects[[1]][i, 2], ")")
      ide = paste0(x-1, '\n')
      for (game in 1:length(scene$games[[1]]$winner)) {
        for (player in 1:length(scene$games[[1]]$team1[[game]])) {
          ide = paste0(ide, scene$games[[1]]$team1[[game]][player])
          if (scene$games[[1]]$team1[[game]][player] == scene$subjects[[1]][i, 1] && game == scene$subjects[[1]][i, 2]) {
            ide = paste0(ide, '? ')
          } else {
            ide = paste0(ide, ' ')
          }
        }
        if (scene$games[[1]]$winner[[game]] == 1) {
          ide = paste0(ide, '>')
        } else {
          ide = paste0(ide, '<')
        }
        for (player in 1:length(scene$games[[1]]$team2[[game]])) {
          ide = paste0(ide, ' ', scene$games[[1]]$team2[[game]][player])
          if (scene$games[[1]]$team2[[game]][player] == scene$subjects[[1]][i, 1] && game == scene$subjects[[1]][i, 2]) {
            ide = paste0(ide, '?')
          } else {
            ide = paste0(ide)
          }
        }
        ide = paste0(ide, '\n')
      }
    } else if (scene$questions[[1]][i] == 2) {
      wIndex = wIndex + 1
      elementToAdd = paste0("'beat", scene$subjects[[1]][i, 1], scene$subjects[[1]][i, 2], "': wouldBeat(", scene$subjects[[1]][i, 1], ", ", scene$subjects[[1]][i, 2], ")")
      ide = paste0(x-1, "-", scene$subjects[[1]][i, 1],",", scene$subjects[[1]][i, 2])
    }
    
    #1	2-1,1	0.231
     else if (scene$questions[[1]][i] == 3) {
      cIndex = cIndex + 1
      elementToAdd = paste0("'close", scene$subjects[[1]][i, 2], "': isClose(", scene$subjects[[1]][i, 1], ", ", scene$subjects[[1]][i, 2], ")")
      ide = paste0(x-1, "-", scene$subjects[[1]][i, 1],",", scene$subjects[[1]][i, 2])
    }
  
    if (i < length(scene$questions[[1]])) {
      elementToAdd = paste0(elementToAdd, ", ")
    }
    print(towModel)
    print(retObject)
    print(elementToAdd)
    retObject = paste0(retObject, elementToAdd)
  
    retObject = paste0(retObject, "}}")
  
    newModel = paste0(towModel, retObject)
  
    dataToWebPPL = list(lazinessPrior = 0.3, 
                        lazyPulling = 0.5,
                        closeThreshold = 10,
                        whodata = whoiswhat,
                        matchInfo = scene)
  
    rs <- webppl(program_code = newModel, 
                 model_var = "tugOfWarModel",
      # inference_opts = list(method = 'rejection', samples = 1000),
      inference_opts = list(method = 'rejection', samples = 100),
      data = dataToWebPPL,
      data_var = "datafromR"
    )
    
    # rs.tidy <- rs %>%
    # gather(key, val, -c(Iteration,Chain))
    # value  of obs
    rs$value <- as.double(rs$value)
    
    if (scene$questions[[1]][i] == 0) {
      df.strength[sIndex,] <- c(ide, mean(rs$value))
    } else if (scene$questions[[1]][i] == 1) {
      df.lazy[lIndex,] <- c(ide, mean(rs$value))
    } else if (scene$questions[[1]][i] == 2) {
      df.wouldwin[wIndex,] <- c(ide, mean(rs$value))
    }
  }
}

print(df.strength)
print(df.lazy)
print(df.wouldwin)
#df.lazy = df.lazy %>% 
#  mutate(rating = as.numeric(rating),
#         rating = 1-rating,
#         trial = 1:nrow(.)) %>% 
#  select(trial,id,rating)

# save(df.lazy,file = 'tow_exp2_prediction.RData')

```

```{r plotResults}

#df.plot = df.results %>% 
#  mutate(rating = rating-50) %>%
#  mutate(id = as.factor(id),
#        winner = factor(winner,levels = c("1","2"),labels = c('Win', "Loss")),
#         rating = ifelse(winner == "Loss", rating*-1, rating))

ggplot(df.strength,aes(x = id, y = as.numeric(rating)))+
  stat_summary(fun.y = 'identity',geom='bar',color = 'black',fill = 'gray80')+
  labs(y = 'strength', x = 'tournament-player')+
  theme_bw()+
  scale_y_continuous(limits = c(0,100))+
  theme(text = element_text(size = 10),
        panel.grid = element_blank())

ggplot(df.lazy, aes(x = reorder(id, as.double(rating)), y = rating))+
  stat_summary(fun.y = 'identity',geom='bar',color = 'black',fill = 'gray80')+
  labs(y = 'probability of laziness', x = 'tournament-player,match')+
  theme_bw()+
  theme(text = element_text(size = 10),
        panel.grid = element_blank())

ggsave('model2_bars.pdf',width=25,height=6)

ggplot(df.wouldwin,aes(x = id, y = rating))+
  stat_summary(fun.y = 'identity',geom='bar',color = 'black',fill = 'gray80')+
  labs(y = 'probability of player 1 win', x = 'tournament-player1,player2')+
  theme_bw()+
  theme(text = element_text(size = 10),
        panel.grid = element_blank())



```

```{r saveResults}

save(list = c("df.results"),file = 'exp1_predictions.RData')

```